---
title: "Practical Machine Learning Course Project Report"
author: "Quan"
date: '2015-12-13'
output: html_document
---
### 1. Introduction    
This project aims at predicting the manner in which ego-surveillance device users did their exercises. The goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
This report adopted an __ensemble model__ for prediction. __Five__ individual classification models are ensembled, namely 1. Linear discriminant analysis (LDA); 2. Neuron network (NNET); 3. Random forest (RF); 4. Linear support vector machine (SVMLin); 5. Radial support vector machine (SVMRad). Pre-process by Principle component analysis was carried out as training control. Variables with over 80% correlation have been collapsed into PCA variables in model training.
The final model is ensembled from the aforementioned models by random forest classification. Final prediction is made and submitted by ensemble model in __Result__ section. The accuracy and __out-of-sample error__ of all models are evaluated in __discussion__ section.

### 2. Codes are Processes.    
#### 2.1 set global seed.    
```{r results="hide", message="FALSE", warning="FALSE"}
set.seed(7900)
```

#### 2.2 quote required packages.    
```{r results="hide", message="FALSE", warning="FALSE"}
library(caret)
library(randomForest)
```

#### 2.3 load assignment data.    
```{r results="hide", message="FALSE", warning="FALSE"}
dataset.train <- read.csv("C:/Users/admin/Documents/pml-training.csv", na.strings=c("", "NA", "#DIV/0!"), header=TRUE, sep="," )
dataset.test <- read.csv("C:/Users/Admin/Documents/pml-testing.csv", na.strings=c("", "NA", "#DIV/0!"), header=TRUE, sep=",")
```

#### 2.4 preprocess on data.    
+ 2.4.1 exclude indexing variables.
```{r results="hide", message="FALSE", warning="FALSE"}
dataset.train <- dataset.train[,-c(1:7)]
dataset.test  <- dataset.test[,-c(1:7)]
```

+ 2.4.2 exclude variables with over 50% void values. These 100 variables are excluded from machine learning process. 
```{r results="hide", message="FALSE", warning="FALSE"}
fun.avecNA <- function(x)  { if(sum(is.na(dataset.train[, x])) > 0.50*nrow(dataset.train)) {return(FALSE)} else {return(TRUE)}  }
var.avecNA <- sapply(X= colnames(dataset.train), FUN=fun.avecNA)
dataset.train <- dataset.train[, var.avecNA]
dataset.test  <- dataset.test[, var.avecNA]
```

+ 2.4.3 preprocess with principle component analysis where correlated variables are compressed.
```{r results="hide", message="FALSE", warning="FALSE"}
var.avecClasse <- colnames(dataset.train) == "classe"
var.correlated <- findCorrelation(cor(dataset.train[, !var.avecClasse]), cutoff=0.8)
names(dataset.train)[var.correlated]
## The existence of highly correlated variables shows the necessity of applying the principle component anlysis.
var.TrainControl <- trainControl(method = "cv", preProcOptions="pca", number = 7, verboseIter=FALSE , allowParallel=TRUE)
```

#### 2.5 devide training data into training and subtesting (ensemble) subsets.    
```{r results="hide", message="FALSE", warning="FALSE"}
subtrain <- createDataPartition(y=dataset.train$classe, p=0.80, list=FALSE)
dataset.subtrain <- dataset.train[subtrain, ] 
dataset.subtest <-  dataset.train[-subtrain,]
```

#### 2.6 supervised machine learning under five designated models.    
```{r results="hide", message="FALSE", warning="FALSE"}
mod.LDA <- train(classe ~ ., data = dataset.subtrain, method = "lda", trControl= var.TrainControl)
mod.NNET <- train(classe ~ ., data = dataset.subtrain, method = "nnet",  verbose=FALSE, trControl= var.TrainControl)
mod.RF <- train(classe ~ ., data = dataset.subtrain, method = "rf", trControl= var.TrainControl)
mod.SVMLin <- train(classe ~ ., data = dataset.subtrain, method = "svmLinear", trControl= var.TrainControl)
mod.SVMRad <- train(classe ~ ., data = dataset.subtrain, method = "svmRadial", trControl= var.TrainControl)
```

#### 2.7 acquire prediction under each of the models on ensemble dataset `dataset.subtest`.    
```{r results="hide", message="FALSE", warning="FALSE"}
pred.LDA     <- predict(object = mod.LDA,    newdata = dataset.subtest)
pred.NNET    <- predict(object = mod.NNET,   newdata = dataset.subtest)
pred.RF      <- predict(object = mod.RF,     newdata = dataset.subtest)
pred.SVMLin  <- predict(object = mod.SVMLin, newdata = dataset.subtest)
pred.SVMRad  <- predict(object = mod.SVMRad, newdata = dataset.subtest)
```

#### 2.8 establish ensemble model.    
```{R results="hide", message="FALSE", warning="FALSE"}
dataset.ensemble <- data.frame(LDA = pred.LDA, NNET = pred.NNET, RF = pred.RF, SVMLin = pred.SVMLin, SVMRad = pred.SVMRad, classe = dataset.subtest$classe)
mod.ENS <- train(classe ~ ., data = dataset.ensemble, method = "rf")
```

#### 2.9 acquire prediction under ensemble model.     
```{r results="hide", message="FALSE", warning="FALSE"}
predt.LDA     <- predict(object = mod.LDA,    newdata = dataset.test)
predt.NNET    <- predict(object = mod.NNET,   newdata = dataset.test)
predt.RF      <- predict(object = mod.RF,     newdata = dataset.test)
predt.SVMLin  <- predict(object = mod.SVMLin, newdata = dataset.test)
predt.SVMRad  <- predict(object = mod.SVMRad, newdata = dataset.test)
dataset.test.ens <- data.frame(LDA = predt.LDA, NNET = predt.NNET, RF = predt.RF, SVMLin = predt.SVMLin, SVMRad = predt.SVMRad)
pred.ENS <- predict(object = mod.ENS, newdata = dataset.test.ens)
```


### 3. Result    
The result of the predition by ensemble model is as follows.
```{r}
pred.ENS
```


### 4. Discussion    
#### 4.1 The accuracy of prediction is improved. Evidence is as follows. Models with multiple solutions retain its optimum.
```{r}
mod.LDA$results$Accuracy
mod.NNET$results$Accuracy
mod.RF$results$Accuracy 
mod.SVMLin$results$Accuracy
mod.SVMRad$results$Accuracy
mod.ENS$results$Accuracy
```

#### 4.2 Out of sample error of each model.
+ __out-of-sample error__ is estimated by examine the prediction quality on subtesting (ensemble) dataset `dataset.subtest`. It can be estimated that an everage error rate of 19/3923 or 0.484% is expected to occur in prediction.
```{r}
mod.ENS$finalModel$confusion
```
